Numerical simulation of the Cordilleran ice sheet through the last glacial cycle

### 2014-06-12 cordillera-cycle.tex
Rework results for structure and avoid repetition.

For all proxy records, we firstly run a low-resolution simulation, using
a~10\,\unit{km}, 300 by 150~points horizontal grid for SSA and SIA velocities,
up to 51~enthalpy layers in the ice, and 31~temperature layers in the bedrock.
For the GRIP and EPICA records, we then run a second simulation using a higher
horizontal resolution of 6\,\unit{km}, up to 101~ice layers, and 61~bedrock
layers.

@@

Despite large differences in the input temperature offset time-series, the
model output presents consistent features that can be observed across the range
of forcing data used. In all simulations, total ice volume is relatively low
during most of the glacial cycle, while two major glaciation events occur
during Marine Isotope Stage (MIS)~4 (61.9--55.4\,\unit{kyr}) and MIS~2
(29.5--16.9\,\unit{kyr}, Fig.~\ref{fig:timeseries}). A local ice minimum is
consistently attained between these two events, during MIS~3
(52.2--45.6\,\unit{kyr}).
However, the magnitude and timing of these events vary significantly
between our simulations.

Simulations forced by the Greenland ice core records (GRIP, NGRIP) produce
highest variability in modelled ice volume throughout the glaciation history.
Although simulations forced by the Antarctic (EPICA, Vostok) and GRIP records
reach the last glacial maximum in ice volume at 19.5--16.9\,\unit{kyr}, those
driven by oceanic (ODP~1012, ODP~1020) and the NGRIP record attain it several
thousands of years earlier. The NGRIP record is the only temperature forcing
that predicts a more extensive glaciation during MIS~4 than during the LGM.

Simulations forced by oceanic and Antarctic records generally result in lower
modelled ice volume during MIS~4 and larger ice volume during MIS~3, thus lower
ice volume variability throughout the simulation length. The ODP~1012
simulation predicts rapid deglaciation prior to 20\,\unit{kyr}
before present due to a local temperature maximum inconsistent with all other
temperature forcing time series used. The ODP~1020 simulation predicts an early
last glacial volume maximum (29.5\,\unit{kyr}), followed by slower deglaciation
than modelled using other records. Finally, simulations forced by Antarctic ice
core records produce a continuous, rapid deglaciation after the last glacial
maximum, whereas the simulation driven by the GRIP record produce a rapid
deglaciation in three steps, including stand-stills and readvances of the ice
margin.

Despite the different timing, snapshots of model output show relatively
consistent patterns of glaciation (Fig.~\ref{fig:snapshots}). As a result of
the different scaling factors applied to palaeo-temperature proxy records, the
ice sheet geometry at the last glacial maximum appears very similar from one
simulation to the next. It includes a central divide located at about
3\,500\,\unit{m} elevation along the spine of the Rocky Mountains. In all
simulations, a central ice cap persists over the Skeena mountains during MIS~3
between the two glaciation events. However, the size of this ice cap, as well
as the magnitude of the MIS~4 glaciation, depends sensitively on the choice of
palaeo-temperature proxy record used to drive the model.

>>>>

For all proxy records, we firstly run a low-resolution simulation, using
a~10\,\unit{km}, 300 by 150~points horizontal grid for SSA and SIA velocities,
up to 51~enthalpy layers in the ice, and 31~temperature layers in the bedrock.
For the GRIP and EPICA records, we then run a second simulation using a higher
horizontal resolution of 6\,\unit{km}, up to 101~ice layers, and 61~bedrock
layers. In the rest of this paper, we refer to different model runs by the
proxy record used for palaeo-temperature forcing.

@@

Despite large differences in input temperature offset time-series, the model
output presents consistent features that can be observed across the range of
forcing data used. In all simulations, total ice volume is relatively low
during most of the glacial cycle, while two major glaciation events occur
during Marine Isotope Stage (MIS)~4 (61.9--55.4\,\unit{kyr}) and MIS~2
(29.5--16.9\,\unit{kyr}, Fig.~\ref{fig:timeseries}). A local ice minimum is
consistently attained during MIS~3 (52.2--45.6\,\unit{kyr}). However, the
magnitude and timing of these events depend significantly on the choice of
proxy record used for palaeo-temperature forcing.

Simulations forced by the Greenland ice core records (GRIP, NGRIP) produce
highest variability in modelled ice volume throughout the glaciation history.
Model runs using oceanic and Antarctic records generally result in lower
modelled ice volume during MIS~4 and larger ice volume during MIS~3, thus
suggesting lower ice volume variability throughout the simulation length. The
NGRIP palaeo-temperature forcing is the only one that results in larger
glaciation during MIS~4 than during the last glacial maximum.

While simulations driven by the Antarctic (EPICA, Vostok) and GRIP records
reach a last glacial maximum in ice volume between 19.5 and 16.9\,\unit{kyr},
those steered by oceanic (ODP~1012, ODP~1020) and the NGRIP record attain it
several thousands of years earlier. Moreover, the ODP~1012 run yields rapid
deglaciation prior to 20\,\unit{kyr} due to a local temperature maximum that is
inconsistent with all other records used. The ODP~1020 simulation predicts an
early last glacial maximum in ice volume at 29.5\,\unit{kyr}, followed by
slower deglaciation than modelled using other records. Finally, model runs
informed by Antarctic ice core records produce a continuous, rapid deglaciation
after the last glacial maximum, whereas the simulations driven by the GRIP
record produce a rapid deglaciation in three steps, including stand-stills and
readvances of the ice margin.

Despite the different timing, snapshots of model output show relatively
consistent patterns of glaciation (Fig.~\ref{fig:snapshots}). As a result of
different scaling factors applied to palaeo-temperature proxy records, the
ice sheet geometry at the last glacial maximum appears very similar from one
simulation to the next. It includes a central divide located at about
3\,500\,\unit{m} elevation along the spine of the Rocky Mountains. In all
simulations, a central ice cap persists over the Skeena mountains during MIS~3
between the two glaciation events. However, the size of this ice cap, as well
as the magnitude of the MIS~4 glaciation, depends sensitively on the choice of
palaeo-temperature proxy record used to drive the model.
